Hello, this material is a didactic study resource in the R language, where I will present some practical knowledge of data visualization using graphs.


Histogram Graph
```{r}

trees
hist(trees$Girth)
hist(trees$Girth, main='Arvores', xlab= 'Altura', ylab= 'Frequencia', col= 'blue')

```

Density Plot
```{r}
densidade = density(trees$Girth)
plot(densidade)
```

Density Plot over Histogram
```{r}
hist(trees$Girth, main = NULL, xlab = NULL, ylab = NULL)
par(new=TRUE)
plot(densidade)
```

Scatter Plot
```{r}
plot(trees$Girth,trees$Height)
plot(trees$Girth,trees$Height, main = 'Arvores', xlab = 'Volume', ylab = 'Circunferencia', col= 'blue', pch=10)
```
```{r}
plot(trees$Girth,trees$Height, main = 'Arvores', xlab = 'Volume', ylab = 'Circunferencia', col= 'blue', pch=10, type = 'b')
```

Correct overlapping points with the same value through jittering
```{r}
plot(jitter(trees$Girth), trees$Height, main = 'Arvores', xlab = 'Volume', ylab = 'Circunferencia', col= 'blue',)
```

Apply legend with categorical dimension
```{r}
plot(CO2$conc, CO2$uptake, pch=20, col=CO2$Treatment)
legend("bottomright",legend=c("nonchilled","chilled"),cex=1,fill=c("black","red"))
```

New data
```{r}
plot(trees)
```

Screen Splitting
```{r}
split.screen(figs = c(2,2))
screen(1)
plot(trees$Girth,trees$Height)
screen(2)
plot(trees$Girth,trees$Volume)
screen(3)
plot(trees$Height,trees$Volume)
screen(4)
hist(trees$Volume)
close.screen(all=TRUE)
```

Boxplot
```{r}
boxplot(trees$Volume,main='Arvore',xlab='Volume')
boxplot(trees$Volume,main='Arvore',xlab='Volume', col='blue', horizontal=TRUE)
boxplot(trees$Volume,main='Arvore',xlab='Volume', col='blue', horizontal=TRUE, outline = F)
```

Additional Options for the Boxplot
```{r}
boxplot(trees$Volume,main='Arvore',xlab='Volume', col='blue', notch = TRUE)
boxplot.stats(trees$Volume)
boxplot(trees)
```

Aggregate Information
```{r}
InsectSprays
agregadospray = aggregate(.~spray,data=InsectSprays,sum)
agregadospray
```
Bar Chart
```{r}
barplot(spray$count, col=gray.colors(6), xlab='Spray',ylab ="Total",names.arg =spray$spray)
box()
```

Pie Chart
```{r}
pie(spray$count, labels = spray$spray, main='Spray', col=c(1:6))

```

Pie Chart with Legend
```{r}
pie(spray$count, labels = NULL, main='Spray', col=c(1:6))
legend('bottomright',legend=spray$spray, cex=1, fill=c(1:6))
```

Tables in LaTeX Format
```{r}
#install.packages("stargazer")
library(stargazer)
stargazer(iris)
```

Tables in HTML Format
```{r}
stargazer(iris, type='html')
```
Tables in Text Format
```{r}
stargazer(iris, type='text')
```
Save to Disk
```{r}
stargazer(women, out='woman.text', summary = FALSE)
```
Lattice Package
```{r}
library(lattice)
```

Boxplot Graph using Lattice
```{r}
bwplot(trees$Volume)
bwplot(trees$Volume, main='Arvores', xlab = 'volume')
```

Histogram Graph using Lattice
```{r}
histogram(trees$Volume, main='Arvore', xlab = 'Volume', aspect= 1, type='percent', nint= 5)

```
```{r}
chickwts
aggregate(chickwts$weight, by= list(chickwts$feed), FUN= sum)
histogram(~weight | feed, data=chickwts)
```

Conditional Scatterplot 
```{r}
CO2
xyplot(CO2$conc ~CO2$uptake | CO2$Type)
xyplot(CO2$conc ~CO2$uptake | CO2$Treatment)
```

Dotplot Graph
```{r}
esoph
dotplot(esoph$alcgp ~esoph$ncontrols | esoph$tobgp, data = esoph )
```

Conditional Density
```{r}
densityplot(~CO2$conc | CO2$Treatment, plot.points=FALSE)
```

3D Plots
```{r}
OrchardSprays
cloud(decrease ~rowpos * colpos, data=OrchardSprays)
cloud(decrease ~rowpos * colpos, ,groups=treatment, data=OrchardSprays)
```

Level Plot
```{r}
trees
levelplot(Girth ~Height * Volume, data=trees)
```

